Exploring the Design Space of Differentially Private Bayesian Inference under Hellinger Distance

### 2019-01-23 ICML2019/main.tex
changes to intro

\begin{abstract}
% Several approaches have been explored in order to make Bayesian inference
% differentially private.
% For instance, \cite{dimitrakakis2014robust} and
% \cite{wang2015privacy} proved that, under specific conditions of the likelihood function and the prior distribution,
% sampling from the posterior distribution, is already differentially private. \cite{zhang2016differential} and \cite{foulds2016theory}
% designed differentially private (d.p.) mechanisms by adding laplacian noise
% to the numeric parameters of the posterior distributions. 
% Accuracy, is usally (\cite{??}) measured by providing upper bounds, holding w.h.p, on the
% $\ell_1$ or $\ell_\infty$-distance between the numeric parameters
% of the released posterior distribution and the actual real ones of posterior distribution.
We explore the design space of differentially private Bayesian
inference mechanisms with accuracy measured in terms of Hellinger
distance over posterior distributions.  
We focus on two discrete models for parametric Bayesian
inference: the Beta-Binomial and the Dirichlet-Multinomial
models. We study two mechanisms based on the 
Laplace perturbation of the parameters of the posterior distribution under
$\ell_1$ norm, and compare them with a discrete mechanism
calibrating noise to a smooth upper bound on the Hellinger distance.
Accuracy is measured through the Hellinger distance between the posterior
distribution released by the different mechanisms and the real one.
We compare the accuracy theoretically and experimentally.
\end{abstract}
 % TODO: replace with your keywords

\section{Introduction}
\label{sec_intro}
%%Data analysis and privacy are important
Data analysis techniques have brought significant improvements in a
variety of applications including applications from the medical,
financial, social, and transportation fields. In order to provide
better services, these applications need large amounts of users data
putting at risk the privacy of individual users contributing their data. 
Differential privacy was proposed a decade ago to address these
situations and it is on its way to become the gold standard in data
privacy.

%%Bayesian inference is a standard tool for data analysis which is
%% also a building block of probabilistic programming languages
Bayesian inference is a standard statistical tool that is useful to
estimate a posterior distribution when a prior distribution is given
and some data is observed.

@@

We focus on two classical models: the Beta-Binomial and the
Dirichlet-Multinomial models. In the Beta-Binomial model the goal
is to estimate an unknown fraction of a population from a sequence
of Bernoulli trials each of which is either 0 or 1 and from an initial
prior belief, a Beta distribution, on that fraction; the trials are assumed to be independent
given the value of the bias. In the Dirichlet-Multinomial model a Multinomial
distribution is used to describe categorical data with more than
two possible outcomes. Similarly, the aim is to learn the various
unknown proportions of a population from a sequence of trials
each of which having more than two possible outcomes. Again,
a prior belief, a Dirichlet distribution, is used to describe the initial knowledge
about those fractions. The posterior distributions on the proportions will be
respectively a Beta distribution and a Dirichlet distribution.

In our work  the posterior distribution is the analysis result we obtained from the data.
While the prior distributions are considerd public information,  publishing the posterior distribution
inferred from a sensitive dataset can leak information about the individuals in the dataset.
In order to guarantee differential privacy and to protect the
individuals' data we can add noise to the posterior before releasing it.
The amount of the noise that we need to introduce
depends on the privacy parameter $\epsilon$ and the sensitivity of the inference to
small changes in the data set. 
Sensitivity can be computed in many different ways based on which metric space
we consider on the output space of the mechanism. In the literature on private Bayesian
inference (e.g. \cite{zhang2016differential,xiao2012bayesian}), it is usually measured on
the vector of numbers parametrizing the output distribution using, e.g. the $\ell_1$ norm.
A more natural approach which we explore here, is to measure sensitivity with respect to
a metric on the space of inferred probability distributions.
Another question is that of how to measure accuracy. Again,
this can be answered in different ways based on the metric
imposed on the output space, and yet again
only in few works in literature (e.g. \cite{zhang2016differential})
distances between probability measures have been used for these purposes.

The question that this work aims at answering is whether
an approach based on probability metrics can improve on
the accuracy of approaches based on metrics over
the numeric parameters of the distributions. 
We will see that in some cases this can happen.

\noindent \textbf{Main contributions.}
What follows is a summary of the main contributions of this work.
\begin{itemize}
	\item We design a new differentially private Bayesian inference mechanism based on the standard exponential mechanism.
    \item A full proof of $(\epsilon, \delta)$-differential privacy of the mechanism is given in the paper. \gp{it's only eps-d.p for now}
	\item We explored two ways to improve the accuracy: 1) calibrating noise to the sensitivity of a metric over distributions
      (e.g. Hellinger distance ($\hellinger$), $f$-divergences, etc\dots). 2) Using a smooth upper bound on the local sensitivity and scale
      the noise to this smooth bound rather than global sensitivity, to improve the mechanism accuracy.
  \item We implemented the new proposed mechanism and other art-of-state mechanisms, comparing their performances in terms of accuracy and privacy.
\end{itemize}

>>>>

\begin{abstract}
% Several approaches have been explored in order to make Bayesian inference
% differentially private.
% For instance, \cite{dimitrakakis2014robust} and
% \cite{wang2015privacy} proved that, under specific conditions of the likelihood function and the prior distribution,
% sampling from the posterior distribution, is already differentially private. \cite{zhang2016differential} and \cite{foulds2016theory}
% designed differentially private (d.p.) mechanisms by adding laplacian noise
% to the numeric parameters of the posterior distributions. 
% Accuracy, is usally (\cite{??}) measured by providing upper bounds, holding w.h.p, on the
% $\ell_1$ or $\ell_\infty$-distance between the numeric parameters
% of the released posterior distribution and the actual real ones of posterior distribution.
We explore the design space of differentially private Bayesian
inference mechanisms with accuracy measured in terms of the Hellinger
distance between posterior distributions.  
We focus on two discrete models for parametric Bayesian
inference: the Beta-Binomial and the Dirichlet-Multinomial
models. We study two mechanisms based on 
Laplace perturbation of the parameters of the posterior distribution under the
$\ell_1$ norm, and compare them with a discrete mechanism
calibrating noise to a smooth upper bound on the Hellinger distance.
Accuracy is measured through the Hellinger distance between the posterior
distribution released by a given mechanism and the one obtained via non-private inference.
We compare the accuracy of our mechanisms theoretically and experimentally.
\end{abstract}
 % TODO: replace with your keywords

\section{Introduction}
\label{sec_intro}
%%Data analysis and privacy are important
Modern data analysis techniques have enabled significant advances in a
variety of applications in medicine, finance, social science, and transportation.
In order to provide better services, these applications need large amounts of users' data,
putting at risk the privacy of individual users contributing their data. 
Differential privacy was proposed a decade ago to address privacy concerns
in these situations, and is now the standard for privacy-preserving data analysis.
 %and it is on its way to become the gold standard in data
%privacy.

%%Bayesian inference is a standard tool for data analysis which is
%% also a building block of probabilistic programming languages
Many statistical applications are driven by Bayesian inference.
This is a standard statistical tool in which a prior distribution is combined
with a sample of observed data in order to estimate a new posterior distribution.
%is a standard statistical tool that is useful to
%estimate a posterior distribution when a prior distribution is given
%and some data is observed.

@@

We focus on inference tasks based on two classical statistical models: the Beta-Binomial and the
Dirichlet-Multinomial models. In the Beta-Binomial model, we are given a
dataset consisting of independent Bernoulli trials with bias $p$. Given an initial
prior belief that $p$ follows a Beta distribution, we may use the dataset to infer a
posterior belief that $p$ follows a Beta distribution with updated parameters. This model
generalizes to the Dirichlet-Multinomial model, in which a multinomial distribution
is used to describe categorial data with more than two possible outcomes. Similarly, the goal
is infer a belief about the proportions of each outcome. Given a prior belief that these proportions
come from a Dirichlet distribution, the posterior belief given multinomial data will
also follow a Dirichlet distribution.

%
%In the Beta-Binomial model the goal
%is to estimate an unknown fraction of a population from a sequence
%of Bernoulli trials each of which is either 0 or 1 and from an initial
%prior belief, a Beta distribution, on that fraction; the trials are assumed to be independent
%given the value of the bias. In the Dirichlet-Multinomial model a Multinomial
%distribution is used to describe categorical data with more than
%two possible outcomes. Similarly, the aim is to learn the various
%unknown proportions of a population from a sequence of trials
%each of which having more than two possible outcomes. Again,
%a prior belief, a Dirichlet distribution, is used to describe the initial knowledge
%about those fractions. The posterior distributions on the proportions will be
%respectively a Beta distribution and a Dirichlet distribution.

In our work, we consider mechanisms which output a complete description of the posterior
distribution inferred using private data.
While the prior distribution associated to a given inference task is considered public information,
the inferred posterior distribution may leak information about the individuals in the dataset.
In order to guarantee differential privacy, we introduce random noise to the posterior before releasing it.
Our goal is to design mechanisms which are accurate, in the sense that this noisy posterior is close
to the posterior one would have inferred without concern for privacy. The most natural way
to measure closeness is via a distance between probability distributions -- nevertheless, only a few works 
in the literature (e.g., \cite{zhang2016differential}) actually state accuracy guarantees in this way.

The amount of the noise that we need to introduce
depends on the differential privacy parameter $\epsilon$ and the sensitivity of the inference to
small changes in the input dataset.
As with accuracy, sensitivity may be measured in many different ways, with each choice giving rise to a different
method of introducing noise for differential privacy.
%Sensitivity can be computed in many different ways based on which metric space
%we consider on the output space of the mechanism. 
In the existing literature on private Bayesian
inference (e.g. \cite{zhang2016differential,xiao2012bayesian}), sensitivity is measured via a norm (e.g., the $\ell_1$ norm)
on the the vector of numbers parametrizing the output distribution. But when the accuracy of a mechanism is determined by
a distance on distributions themselves, this may not be such a natural choice. The distance between
two parameterized distributions may behave wildly as a function of the distance between the vectors of parameters characterizing them.

A more natural approach, which we propose here, is to also directly measure sensitivity with respect
to a metric on the space of inferred probability distributions. Our goal is to identify cases where jointly measuring
accuracy and sensitivity in terms of a single metric on probability distributions can yield improved accuracy. 
{\color{red}{MB: Is this accurate? Would we ever want to use different metrics for accuracy and sensitivity?}}

%A more natural approach which we explore here, is to measure sensitivity with respect to
%a metric on the space of inferred probability distributions.
%Another question is that of how to measure accuracy. Again,
%this can be answered in different ways based on the metric
%imposed on the output space, and yet again
%only in few works in literature (e.g. \cite{zhang2016differential})
%distances between probability measures have been used for these purposes.

%The question that this work aims at answering is whether
%an approach based on probability metrics can improve on
%the accuracy of approaches based on metrics over
%the numeric parameters of the distributions. 
%We will see that in some cases this can happen.

\noindent \textbf{Main contributions.}
What follows is a summary of the main contributions of this work.
\begin{itemize}
	\item We give a new differentially private Bayesian inference mechanism based on the standard exponential mechanism.
   %    \item A full proof of $(\epsilon, \delta)$-differential privacy of the mechanism is given in the paper. 
	\item We propose two ways of improving the accuracy of this basic mechanism: 1) We calibrate noise to a measure of sensitivity determined by a metric between distributions
      (e.g. Hellinger distance ($\hellinger$), $f$-divergences, etc\dots). 2) We scale noise according to a a smooth upper bound on the local sensitivity, rather than according to global sensitivity.
  \item We implement the new proposed mechanism and other art-of-state mechanisms, comparing their performances in terms of accuracy and privacy.
\end{itemize}
